---
title: "Rhopilema single-cell analysis"
author: "Matéo Peyroche"
format: html
---

#DATA ANLYSIS OF THE UPDATED RHOPILEMA DATASET

##Importing libraries, dataset and relevant objects

Libraries
```{r}
#| echo: false
set.seed(213)
library(Seurat,quietly=T)
library(SeuratObject,quietly=T)
library(RColorBrewer,quietly=T)
library(Matrix,quietly=T)
library(ggplot2,quietly=T)
library(Polychrome,quietly=T)
library(dplyr,quietly=T)
library(presto,quietly=T)
library(rtracklayer,quietly=T)
library(stringr,quietly=T)
library(tidyverse,quietly=T)
library(colorspace,quietly=T)
library(patchwork,quietly=T)
library(scCustomize,quietly=T)


```

Loading the dataset and the gene annotations
```{r}
load (file= "../../data/scyphozoa/RhAllData.RData")
gene_table.rh <- read.table(file="../../data/seqs/genomes/RhoEs_GCF_013076305.1/RhoEs_all_protein_gene_table.tsv", header=F)
names(gene_table.rh)[names(gene_table.rh)=="V1"] <- "protein_id"
names(gene_table.rh)[names(gene_table.rh)=="V2"] <- "gene_id"

data.rh <- Rh.alldata
data.rh <- JoinLayers(data.rh)
rh.ephyra <- subset(data.rh, subset = orig.ident == "Rh.ephyra")

out_dir <- '../../screenshots/R/Rh/final/' #output directory for figures

```

#Visualization of the dataset

```{r}

#set IDs to idents
Idents(data.rh) <- data.rh$IDs

#match the idents with the colors of the palette
coarse_cluster_levels <- levels(Idents(data.rh))


coarse_pal <- createPalette(
  seedcolors = c("#EE7733", "#009988", "#EE3377"),
  N = length(coarse_cluster_levels),
  range = c(20,90)
)

names(coarse_pal) <- coarse_cluster_levels
coarse_pal <- sortByHue(coarse_pal)

#print umap
rh_dimplot_coarse <- DimPlot(
  data.rh,
  reduction = "umap",
  cols = coarse_pal,
  pt.size = 0.5
)

DimPlot(
  data.rh,
  reduction = "umap",
  group.by = "IDs",     # what determines the color
  split.by = "orig.ident",       # what determines the panel split
  cols = coarse_pal,
  pt.size = 0.5
)


Idents(data.rh) <- data.rh$ID.separate
Idents(rh.ephyra) <- rh.ephyra$ID.separate


#match the idents with the colors of the palette
fine_cluster_levels <- levels(Idents(data.rh))
fine_pal <- character(length(fine_cluster_levels))
names(fine_pal) <- fine_cluster_levels

tab <- table(fine = data.rh@meta.data$ID.separate, coarse = data.rh@meta.data$IDs)

# for each fine cluster (row), pick the coarse cluster (column) with the most cells =
#pair fine clusters to the coarse cluster they are derived from
fine_to_coarse <- apply(tab, 1, function(x) names(which.max(x)))

#Derive fine cluster color from the coarse one

for (coarse_id in names(coarse_pal)) {
  children <- names(fine_to_coarse)[fine_to_coarse == coarse_id]
  n_children <- length(children)
  base_col <- coarse_pal[coarse_id]

  if (n_children == 1) {
    # no subclustering -> keep the coarse color exactly as-is
    fine_pal[children] <- base_col
  } else {
    # spread shades from darker to lighter around the base color
    shades <- sapply(seq(-0.4, 0.4, length.out = n_children), function(amt) {
      if (amt < 0) darken(base_col, abs(amt)) else lighten(base_col, amt)
    })
    fine_pal[children] <- shades
  }
}

rh_dimplot_fine <-DimPlot(
  data.rh,
  reduction = "umap",
  cols = fine_pal,
  pt.size = 0.5
)+ guides(color = guide_legend(ncol = 1, override.aes = list(size = 3)))

rh_dimplot_fine_p_v_e <- DimPlot(
  data.rh,
  reduction = "umap",
  group.by = "ID.separate",     # what determines the color
  split.by = "orig.ident",       # what determines the panel split
  cols = fine_pal,
  pt.size = 0.5
)+ guides(color = guide_legend(ncol = 1, override.aes = list(size = 3)))

ggsave(
  filename = file.path(out_dir, "rh_dimplot_fine_p_v_e2.png"),
  plot = rh_dimplot_fine_p_v_e,
  width = 12,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "rh_dimplot_fine2.png"),
  plot = rh_dimplot_fine,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "rh_dimplot_coarse2.png"),
  plot = rh_dimplot_coarse,
  width = 10,
  height = 6,
  dpi = 300
)


```


Highlighting the muscle cluster in the all and ephyra datasets 
```{r}

rh_highlight_plot_ST_all <- Cluster_Highlight_Plot(
  seurat_object = data.rh,
  reduction = "umap",
  cluster_name = "20_12.muscle.ST",
  highlight_color = "gold",
  background_color = "lightgray"
)

rh_highlight_plot_ST_eph <- Cluster_Highlight_Plot(
  seurat_object = rh.ephyra,
  reduction = "umap",
  cluster_name = "20_12.muscle.ST",
  highlight_color = "gold",
  background_color = "lightgray"
)

rh_highlight_plot_ST_all
rh_highlight_plot_ST_eph

ggsave(
  filename = file.path(out_dir, "rh_highlight_plot_ST_all.png"),
  plot = rh_highlight_plot_ST_all,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "rh_highlight_plot_ST_eph.png"),
  plot = rh_highlight_plot_ST_eph,
  width = 10,
  height = 6,
  dpi = 300
)

```


Creating gene labels for dotplots by transferring Aurelia names from the OMA repaired table, or the NCBI annotations. 
```{r}
OMA.ReAc <- read_tsv(file = "../../data/OMA/re2-kostya_repaired.tsv")
load (file= "../../data/scyphozoa_AGC/ACOE.genes.RData")
ac_to_rh.names <- OMA.ReAc %>%
  left_join(genes.ac %>% select(ACO, gene_short_name, cellranger), by = c('Gene.Ac'='cellranger'))

#Transferring names from Aurelia based on the OMA map 
#When 1:1, or many:1 the Aurelia name is directly passed, in 1:many, Aurelia names are fused
#This is done on anything that is not an ACOE name. 
ac_to_rh.names <- ac_to_rh.names %>%
  filter(pairing %in% c('1:1','many:1','1:many'))
pattern <- "^[A-Z]+[0-9]+G[0-9]+$"
ac_to_rh.names <- ac_to_rh.names %>%
  filter(!grepl(pattern, gene_short_name))
ac_to_rh.names <- ac_to_rh.names %>%
  group_by(Gene.Re) %>%
  summarise(
    gene_short_name = paste(gene_short_name, collapse = "_")
  ) %>%
  ungroup()

ac_to_rh.names<- genes.Rh %>%
  left_join(ac_to_rh.names, by = c('gene_id'='Gene.Re'))

ac_to_rh.names$product <- gsub("uncharacterized ","", ac_to_rh.names$product)

ac_to_rh.names <- ac_to_rh.names %>%
  mutate(
    name = coalesce(
      na_if(gene_short_name, "-"),
      na_if(product, "-"),
      na_if(gene_id, "-")
    )
  )

gene_labels <- ac_to_rh.names %>%
  group_by(gene_id) %>%
  summarise(label = dplyr::first(name), .groups = "drop") %>%
  mutate(label = ifelse(nchar(label) > 20, paste0(substr(label, 1, 20), "..."), label)) %>%
  tibble::deframe()

```


Plotting of orthologs of Aurelia striated muscle specific genes
```{r}
#Plotting stmyhc1 and 4 and the "calm-like-37" orthologs
rh_dotplot_stm_struct_genes.e <- DotPlot(rh.ephyra, features = c('LOC135691812','LOC135691803','LOC135688481')) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = gene_labels[c('LOC135691812','LOC135691803','LOC135688481')])

rh_featplot_stm_stmyhc.e <- FeaturePlot(rh.ephyra, features = c('LOC135691812','LOC135691803','LOC135688481'),reduction = "umap")&
  scale_color_gradientn(colors = c("lightgrey", "#FDBB84", "#B30000"))

rh_dotplot_stm_struct_genes.e
rh_featplot_stm_stmyhc.e

rh_dotplot_stm_struct_genes.all <- DotPlot(data.rh, features = c('LOC135691812','LOC135691803','LOC135688481')) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = gene_labels[c('LOC135691812','LOC135691803','LOC135688481')])

rh_featplot_stm_stmyhc.all <- FeaturePlot(data.rh, features = c('LOC135691812','LOC135691803','LOC135688481'),reduction = "umap")&
  scale_color_gradientn(colors = c("lightgrey", "#FDBB84", "#B30000"))

rh_dotplot_stmyhc4_p_v_e <- DotPlot(data.rh, features = c('LOC135691803'), split.by = "orig.ident", cols = c('#2166AC','#B30000'))+  scale_x_discrete(labels = gene_labels[c('LOC135691803')])

rh_dotplot_stm_struct_genes.all
rh_featplot_stm_stmyhc.all
rh_dotplot_stmyhc4_p_v_e

ggsave(
  filename = file.path(out_dir, "rh_dotplot_stm_struct_genes_e.png"),
  plot = rh_dotplot_stm_struct_genes.e,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "rh_featplot_stm_stmyhc_e.png"),
  plot = rh_featplot_stm_stmyhc.e,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "rh_dotplot_stm_struct_genes_all.png"),
  plot = rh_dotplot_stm_struct_genes.all,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "rh_featplot_stm_stmyhc_all.png"),
  plot = rh_featplot_stm_stmyhc.all,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "rh_dotplot_stmyhc4_p_v_e.png"),
  plot = rh_dotplot_stmyhc4_p_v_e,
  width = 8,
  height = 8,
  dpi = 300
)


```

#Unbiased approach

##Ephyra Differentially expressed genes (DEG)

As we are interested in the striated muscles, we will work on the ephyra. This is especially relevant here as the STmuscle cluster seems to also include polyp contractile cells, but the Aurelia ST-muscle specific genes from Aurelia seem to only be expressed in the subcluster kept in the ephyra subset. 
Getting DEGs for both clustering and for the TF genes subset 
```{r}
#Getting the transcription factor list from the AnimalTFDB4.0 TF Predict output

Rh_TF_table <- read.csv(file = '../../data/seqs/TF_lists/RhoEs_TF_list.csv')
Rh_TF_table <- Rh_TF_table %>%
  left_join(gene_table.rh, 
             by = c('TF'='protein_id'))
Rh_TF_table <- Rh_TF_table %>%
  distinct(gene_id, .keep_all = T)
Rh_TF_list <- Rh_TF_table$gene_id
 
#First for the coarse clustering

Idents(rh.ephyra) <- rh.ephyra$IDs

rh.all.markers_coarse.e <- FindAllMarkers(
  rh.ephyra,
  assay = "RNA",
  logfc.threshold = 0.25,
  return.thresh   = 0.001,
  min.pct         = 0.2,
  only.pos        = TRUE
)

rh.tf.markers_coarse.e <- FindAllMarkers(
  rh.ephyra,
  features= intersect(Rh_TF_list,rownames(rh.ephyra)),
  assay = "RNA",
  logfc.threshold = 0.25,
  return.thresh   = 0.001,
  min.pct         = 0.06,
  only.pos        = TRUE
)


#For the fine clustering
Idents(rh.ephyra) <- rh.ephyra$ID.separate

rh.all.markers_fine.e <- FindAllMarkers(
  rh.ephyra,
  assay = "RNA",
  logfc.threshold = 0.25,
  return.thresh   = 0.001,
  min.pct         = 0.2,
  only.pos        = TRUE
)

rh.tf.markers_fine.e <- FindAllMarkers(
  rh.ephyra,
  features= intersect(Rh_TF_list,rownames(rh.ephyra)),
  assay = "RNA",
  logfc.threshold = 0.25,
  return.thresh   = 0.001,
  min.pct         = 0.06,
  only.pos        = TRUE
)
```

##Striated muscle cluster
Extracting the gene markers for the striated muscle
```{r}
rh.muscleST_markers.e <- rh.all.markers_coarse.e %>%
  filter(cluster == "12.muscle.ST") %>%
  arrange(p_val_adj,desc(avg_log2FC))

rh.tf.muscleST_markers.e <- rh.tf.markers_coarse.e %>%
  filter(cluster == "12.muscle.ST") %>%
  arrange(p_val_adj,desc(avg_log2FC))

rh.muscleST_markers.e_fine <- rh.all.markers_fine.e %>%
  filter(cluster == "20_12.muscle.ST") %>%
  arrange(p_val_adj,desc(avg_log2FC))

rh.tf.muscleST_markers.e_fine <- rh.tf.markers_fine.e %>%
  filter(cluster == "20_12.muscle.ST") %>%
  arrange(p_val_adj,desc(avg_log2FC))

#It is virtually exactly the same

```


Plotting the differentially expressed transcription factors.
```{r}
rh_dotplot_STm_tf_markers_all <- DotPlot(rh.ephyra, features = rh.tf.muscleST_markers.e$gene) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = gene_labels[rh.tf.muscleST_markers.e$gene])+labs(title = "Differentially expressed transcription factors of the striated muscles of Rhopilema",size = 20, face = "bold")+
  theme(plot.title = element_text(hjust = 0.5))

print(rh_dotplot_STm_tf_markers_all)

ggsave(
  filename = file.path(out_dir, "rh_dotplot_STm_tf_markers_all.png"),
  plot = rh_dotplot_STm_tf_markers_all,
  width = 10,
  height = 6,
  dpi = 300
)

#Removing the widely expressed genes 
rh_dotplot_STm_tf_markers_filtered <- DotPlot(rh.ephyra, features = rh.tf.muscleST_markers.e$gene[-c(4,6,9)]) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = gene_labels[rh.tf.muscleST_markers.e$gene[-c(4,6,9)]])+
 scale_x_discrete(labels = gene_labels[rh.tf.muscleST_markers.e$gene])+labs(title = "Top differentially expressed transcription factors of the striated muscles of Rhopilema",size = 20, face = "bold")+
  theme(plot.title = element_text(hjust = 0.5))+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))


#Getting the stMuscle subset to show TF pattern expression
rh.ST <- subset(rh.ephyra, idents = "20_12.muscle.ST")

rh_dotplot_STm_tf_markers_filtered

rh_featplot_STmcluster_TFmarkers <- FeaturePlot(rh.ST, features = rh.tf.muscleST_markers.e$gene[-c(4,6,9)],reduction = "umap")&scale_color_gradientn(colors = c("lightgrey", "#FDBB84", "#B30000"))

ggsave(
  filename = file.path(out_dir, "rh_dotplot_STm_tf_markers_filtered.png"),
  plot = rh_dotplot_STm_tf_markers_filtered,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "rh_featplot_STmcluster_TFmarkers.png"),
  plot = rh_featplot_STmcluster_TFmarkers,
  width = 10,
  height = 6,
  dpi = 300
)

```


###Individual look at genes of interest

```{r}
#Otx genes 
#Importing the orthology table from the tree. It includes the uncharacterized homeobox that sits in the cluster
full_otx_ortho_table <- read_tsv("../../synteny/ortho_table_otx_synteny.tsv")
rh.otx_genes <- full_otx_ortho_table %>%
  select(orthogroup,Rhopilema_esculentum)%>%
  filter (!is.na(Rhopilema_esculentum))
rh.otx_genes$orthogroup[6] <- "unch_homeo"
rh.otx_only_genes <- rh.otx_genes[-6,]
rh.unch_home_gene <- rh.otx_genes[6,]

rh_dotplot_otx_cluster.e <- DotPlot(rh.ephyra, features =rh.otx_genes$Rhopilema_esculentum ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = rh.otx_genes$orthogroup)+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))

rh_dotplot_otx_only.e <- DotPlot(rh.ephyra, features =rh.otx_only_genes$Rhopilema_esculentum ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = rh.otx_only_genes$orthogroup)+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))


rh_dotplot_unch_homeo.e <-DotPlot(rh.ephyra, features =rh.unch_home_gene$Rhopilema_esculentum ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = rh.unch_home_gene$orthogroup)+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))

rh_dotplot_otx_cluster.e
rh_dotplot_otx_only.e
rh_dotplot_unch_homeo.e

ggsave(
  filename = file.path(out_dir, "rh_dotplot_otx_cluster.png"),
  plot = rh_dotplot_otx_cluster.e,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "rh_dotplot_otx_only.png"),
  plot = rh_dotplot_otx_only.e,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "rh_dotplot_unch_homeo.png"),
  plot = rh_dotplot_unch_homeo.e,
  width = 10,
  height = 6,
  dpi = 300
)


```


```{r}
#CPATH
rh_dotplot_cpath.e <-DotPlot(rh.ephyra, features ="LOC135682280" ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = gene_labels["LOC135682280"])

rh_dotplot_cpath.e

ggsave(
  filename = file.path(out_dir, "rh_dotplot_cpath.png"),
  plot = rh_dotplot_cpath.e,
  width = 10,
  height = 6,
  dpi = 300
)

```

```{r}
#Hox1
rh_dotplot_hox1.e <-DotPlot(rh.ephyra, features ="LOC135682134")+
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis()+scale_x_discrete(labels = gene_labels["LOC135682134"])

rh_dotplot_hox1.e

ggsave(
  filename = file.path(out_dir, "rh_dotplot_hox1.png"),
  plot = rh_dotplot_hox1.e,
  width = 10,
  height = 6,
  dpi = 300
)
```

```{r}
foxb_genes <- c( 'LOC135691975', 'LOC135692080')

rh_dotplot_foxb.e <-DotPlot(rh.ephyra, features =foxb_genes)+
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis()+scale_x_discrete(labels = c('foxB1 chimera','foxB2'))

rh_dotplot_foxb.e

ggsave(
  filename = file.path(out_dir, "rh_dotplot_foxb.png"),
  plot = rh_dotplot_foxb.e,
  width = 10,
  height = 6,
  dpi = 300
)


```

```{r}
#btbd2
rh_dotplot_btb <- DotPlot(rh.ephyra, features =c("LOC135684936")) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis()+scale_x_discrete(labels = c('btbd2-like-11'))

rh_dotplot_btb

ggsave(
  filename = file.path(out_dir, "rh_dotplot_btb.png"),
  plot = rh_dotplot_btb,
  width = 12,
  height = 14,
  dpi = 300
)

```

Showing that these candidates are not expressed in the polyp
```{r}
rh_dotplot_tf_candi_lifehistory <- DotPlot(data.rh, features = c("LOC135682280", "LOC135681476","LOC135682134","LOC135684936"), group.by = "orig.ident")+
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis()+scale_x_discrete(labels = gene_labels[c("LOC135682280", "LOC135681476","LOC135682134","LOC135684936")])+
     theme(axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ), axis.text.y =element_text(size=15, face="bold" ))


rh_dotplot_tf_candi_lifehistory

ggsave(
  filename = file.path(out_dir, "rh_dotplot_tf_candi_lifehistory.png"),
  plot = rh_dotplot_tf_candi_lifehistory,
  width = 8,
  height = 6,
  dpi = 300
)
```

#Biased approach

Plotting the biased approach transcription factors by importing the built orthology table
```{r}
#Importing the table based on the OMA table and OrthoSearch script 
bilat_tf_table <- read.table(file = "../../data/seqs/bilat_tfs/bilat_tf_orthology_table.tsv",header = T,fill = T)

#removing hlxb9 since it is not hlxB9 (not present in medusozoan)+all genes for which 
#no litterature could be found
filter_out <- c('hlxb9','HOX1','FEZF1','OTP','RAX','EMX')

bilat_tf_table <- bilat_tf_table[!(bilat_tf_table$gene %in% filter_out),] 
bilat_tf_rh <- bilat_tf_table$Rhopilema_gene


# Build the named vector: names = original, values = display labels
bilat_tf_rh_label <- setNames(bilat_tf_table$gene, bilat_tf_table$Rhopilema_gene)


bilat_tf_rh_plot <- DotPlot(
  rh.ephyra,
  features = bilat_tf_rh
) +
  RotatedAxis() +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  scale_x_discrete(labels = bilat_tf_rh_label)+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(angle = 90, vjust = 0.5, hjust = 1))
bilat_tf_rh_plot

ggsave(
  filename = file.path(out_dir, "rh_bilat_tfs_updated_dotplot.png"),
  plot = bilat_tf_rh_plot,
  width = 10,
  height = 6,
  dpi = 300
)

```

Saving the marker genes tables as objects and tsv
```{r}

write.table(rh.tf.muscleST_markers.e, file = "../../results/R/expressed_tfs_lists/rh_ST_expressed_tf.tsv", sep = "\t",  row.names = FALSE, quote = FALSE)
write.table(rh.tf.markers_fine.e, file = "../../results/R/expressed_tfs_lists/rh_tf_markers_fine_clusters.tsv", sep = "\t",  row.names = FALSE, quote = FALSE)
write.table(rh.all.markers_fine.e, file = "../../results/R/expressed_tfs_lists/rh_all_markers_fine_clusters.tsv", sep = "\t",  row.names = FALSE, quote = FALSE)

save(rh.all.markers_coarse.e, rh.all.markers_fine.e, rh.tf.markers_coarse.e, rh.tf.markers_fine.e, file ="../../data/scyphozoa/rh_markers.RData")

```


