---
title: "Sanderia single-cell analysis"
author: "Matéo Peyroche"
format: html
---

#DATA ANLYSIS OF THE UPDATED SANDERIA DATASET

##Importing libraries, dataset and relevant objects

Libraries
```{r}
#| echo: false
set.seed(213)
library(Seurat,quietly=T)
library(SeuratObject,quietly=T)
library(RColorBrewer,quietly=T)
library(Matrix,quietly=T)
library(ggplot2,quietly=T)
library(Polychrome,quietly=T)
library(dplyr,quietly=T)
library(presto,quietly=T)
library(rtracklayer,quietly=T)
library(stringr,quietly=T)
library(tidyverse,quietly=T)
library(colorspace,quietly=T)
library(patchwork,quietly=T)
library(scCustomize,quietly=T)


```

Loading the dataset and the gene annotations
```{r}
load (file= "../../data/scyphozoa/SaAllData.RData")
genes.sa <- read_tsv(file = "../../sanderia/annotations/sanderia_func_annotations.tsv")
gene_table.sa <- read.table(file="../../data/seqs/genomes/Sanderia/Sanma_gene_transcript_protein_table.tsv", header=F)
names(gene_table.sa)[names(gene_table.sa)=="V1"] <- "gene_id"
names(gene_table.sa)[names(gene_table.sa)=="V2"] <- "transcript_id"
names(gene_table.sa)[names(gene_table.sa)=="V3"] <- "protein_id"

data.sa <- Sa.alldata
data.sa <- JoinLayers(data.sa)
sa.ephyra <- subset(data.sa, subset = orig.ident == "Sa.ephyra")

out_dir <- '../../screenshots/R/Sa/final/' #output directory for figures

```


#Visualization of the dataset

```{r}

#set IDs to idents
Idents(data.sa) <- data.sa$IDs

#match the idents with the colors of the palette
coarse_cluster_levels <- levels(Idents(data.sa))


coarse_pal <- createPalette(
  seedcolors = c("#EE7733", "#009988", "#EE3377"),
  N = length(coarse_cluster_levels),
  range = c(20,90)
)

names(coarse_pal) <- coarse_cluster_levels
coarse_pal <- sortByHue(coarse_pal)

#print umap
sa_dimplot_coarse <- DimPlot(
  data.sa,
  reduction = "umap.h",
  cols = coarse_pal,
  pt.size = 0.5
)

Idents(data.sa) <- data.sa$ID.separate

#match the idents with the colors of the palette
fine_cluster_levels <- levels(Idents(data.sa))
fine_pal <- character(length(fine_cluster_levels))
names(fine_pal) <- fine_cluster_levels

tab <- table(fine = data.sa@meta.data$ID.separate, coarse = data.sa@meta.data$IDs)

# for each fine cluster (row), pick the coarse cluster (column) with the most cells =
#pair fine clusters to the coarse cluster they are derived from
fine_to_coarse <- apply(tab, 1, function(x) names(which.max(x)))

#Derive fine cluster color from the coarse one

for (coarse_id in names(coarse_pal)) {
  children <- names(fine_to_coarse)[fine_to_coarse == coarse_id]
  n_children <- length(children)
  base_col <- coarse_pal[coarse_id]

  if (n_children == 1) {
    # no subclustering -> keep the coarse color exactly as-is
    fine_pal[children] <- base_col
  } else {
    # spread shades from darker to lighter around the base color
    shades <- sapply(seq(-0.4, 0.4, length.out = n_children), function(amt) {
      if (amt < 0) darken(base_col, abs(amt)) else lighten(base_col, amt)
    })
    fine_pal[children] <- shades
  }
}

sa_dimplot_fine <- DimPlot(
  data.sa,
  reduction = "umap.h",
  cols = fine_pal,
  pt.size = 0.5
)+ guides(color = guide_legend(ncol = 1, override.aes = list(size = 3)))


#Plotting according to original identity : polyp or ephyra. 
sa_dimplot_fine_p_v_e <- DimPlot(
  data.sa,
  reduction = "umap.h",
  group.by = "ID.separate",     # what determines the color
  split.by = "orig.ident",       # what determines the panel split
  cols = fine_pal,
  pt.size = 0.5
)+ guides(color = guide_legend(ncol = 1, override.aes = list(size = 3)))

print(sa_dimplot_coarse)
print(sa_dimplot_fine)
print(sa_dimplot_fine_p_v_e)

```


Highlighting the muscle cluster in the all and ephyra datasets 
```{r}

sa_highlight_plot_ST_all <- Cluster_Highlight_Plot(
  seurat_object = data.sa,
  reduction = "umap.h",
  cluster_name = "6_6.STmuscle",
  highlight_color = "gold",
  background_color = "lightgray"
)+NoAxes()+labs(title = "Striated muscle cluster of the Sanderia all dataset",size = 20, face = "bold")+
  theme(plot.title = element_text(hjust = 0.5))

sa_highlight_plot_ST_eph <- Cluster_Highlight_Plot(
  seurat_object = sa.ephyra,
  reduction = "umap.h",
  cluster_name = "6_6.STmuscle",
  highlight_color = "gold",
  background_color = "lightgray"
)+NoAxes()+labs(title = "Striated muscle cluster of the Sanderia ephyra dataset",size = 20, face = "bold")+
  theme(plot.title = element_text(hjust = 0.5))

sa_highlight_plot_ST_all
sa_highlight_plot_ST_eph

ggsave(
  filename = file.path(out_dir, "sa_highlight_plot_ST_all.png"),
  plot = sa_highlight_plot_ST_all,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "sa_highlight_plot_ST_eph.png"),
  plot = sa_highlight_plot_ST_eph,
  width = 10,
  height = 6,
  dpi = 300
)

```

Creating gene labels for dotplots by importing names from the annotation file
```{r}
gene_labels <- genes.sa %>%
  group_by(gene_id) %>%
  summarise(label = dplyr::first(name), .groups = "drop") %>%
  mutate(label = ifelse(nchar(label) > 20, paste0(substr(label, 1, 20), "..."), label)) %>%
  tibble::deframe()
```

Plotting of orthologs of Aurelia striated muscle specific genes
```{r}
#Plotting stmyhc1 and 4 and the two "calm-like-37" orthologs
sa_dotplot_stm_struct_genes <- DotPlot(sa.ephyra, features = c('ENSXQXG00000009573','ENSXQXG00000009750','ENSXQXG00000016976','ENSXQXG00000003661')) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = c('stmyhc1','stmyhc4','calm-like-37.1','calm-like-37.2'))

sa_featplot_stm_stmyhc <- FeaturePlot(sa.ephyra, features = c('ENSXQXG00000009573','ENSXQXG00000009750'),reduction = "umap.h")+labs(title = "Expression pattern of stmyhc1 and stmyhc4 in the Sanderia ephyra dataset",size = 20, face = "bold")+
  theme(plot.title = element_text(hjust = 0.5))&
  scale_color_gradientn(colors = c("lightgrey", "#FDBB84", "#B30000"))

sa_dotplot_stm_struct_genes
sa_featplot_stm_stmyhc

ggsave(
  filename = file.path(out_dir, "sa_dotplot_stm_struct_genes.png"),
  plot = sa_dotplot_stm_struct_genes,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "sa_featplot_stm_stmyhc.png"),
  plot = sa_featplot_stm_stmyhc,
  width = 14,
  height = 8,
  dpi = 300
)


```


#Unbiased approach

##Ephyra Differentially expressed genes (DEG)

As we are interested in the striated muscles, we will work on the ephyra
Getting DEGs for both clustering and for the TF genes subset 
```{r}
#Getting the transcription factor list from the annotation file
sa.TFs <- genes.sa %>%
  select(gene_id,TF_Fam)%>%
  filter(!is.na(TF_Fam))

#We will now only work on the ephyra subset 

#First for the coarse clustering

Idents(sa.ephyra) <- sa.ephyra$IDs

sa.all.markers_coarse.e <- FindAllMarkers(
  sa.ephyra,
  assay = "RNA",
  logfc.threshold = 0.25,
  return.thresh   = 0.001,
  min.pct         = 0.2,
  only.pos        = TRUE
)

sa.tf.markers_coarse.e <- FindAllMarkers(
  sa.ephyra,
  features= intersect(sa.TFs$gene_id,rownames(data.sa)),
  assay = "RNA",
  logfc.threshold = 0.25,
  return.thresh   = 0.001,
  min.pct         = 0.06,
  only.pos        = TRUE
)


#For the fine clustering
Idents(sa.ephyra) <- sa.ephyra$ID.separate

sa.all.markers_fine.e <- FindAllMarkers(
  sa.ephyra,
  assay = "RNA",
  logfc.threshold = 0.25,
  return.thresh   = 0.001,
  min.pct         = 0.2,
  only.pos        = TRUE
)

sa.tf.markers_fine.e <- FindAllMarkers(
  sa.ephyra,
  features= intersect(sa.TFs$gene_id,rownames(data.sa)),
  assay = "RNA",
  logfc.threshold = 0.25,
  return.thresh   = 0.001,
  min.pct         = 0.06,
  only.pos        = TRUE
)
```


##Striated muscle cluster
Extracting the gene markers for the striated muscle
```{r}
sa.muscleST_markers.e <- sa.all.markers_coarse.e %>%
  filter(cluster == "6.STmuscle") %>%
  arrange(p_val_adj,desc(avg_log2FC))

sa.tf.muscleST_markers.e <- sa.tf.markers_coarse.e %>%
  filter(cluster == "6.STmuscle") %>%
  arrange(p_val_adj,desc(avg_log2FC))

```


Plotting the differentially expressed transcription factors.
```{r}
sa_dotplot_STm_tf_markers_all <- DotPlot(sa.ephyra, features = sa.tf.muscleST_markers.e$gene) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = gene_labels[sa.tf.muscleST_markers.e$gene])

print(sa_dotplot_STm_tf_markers_all)

sa_dotplot_STm_tf_markers_top10 <- DotPlot(sa.ephyra, features = sa.tf.muscleST_markers.e[1:10,]$gene) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = gene_labels[sa.tf.muscleST_markers.e[1:10,]$gene])+labs(title = "Top differentially expressed transcription factors of the striated muscles of Sanderia",size = 20, face = "bold")+
  theme(plot.title = element_text(hjust = 0.5))+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))

print(sa_dotplot_STm_tf_markers_top10)

ggsave(
  filename = file.path(out_dir, "sa_dotplot_STm_tf_markers_all.png"),
  plot = sa_dotplot_STm_tf_markers_all,
  width = 14,
  height = 8,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "sa_dotplot_STm_tf_markers_top10.png"),
  plot = sa_dotplot_STm_tf_markers_top10,
  width = 10,
  height = 6,
  dpi = 300
)


```

###Individual look at genes of interest

```{r}
#Otx genes 
#Importing the orthology table from the tree. It includes the uncharacterized homeobox that sits in the cluster
full_otx_ortho_table <- read_tsv("../../synteny/ortho_table_otx_synteny.tsv")
sa.otx_genes <- full_otx_ortho_table %>%
  select(orthogroup,Sanderia_malayensis)%>%
  filter (!is.na(Sanderia_malayensis))
sa.otx_genes$orthogroup[6] <- "unch_homeo"
sa.otx_only_genes <- sa.otx_genes[-6,]
sa.unch_home_gene <- sa.otx_genes[6,]

sa_dotplot_otx_cluster.e <- DotPlot(sa.ephyra, features =sa.otx_genes$Sanderia_malayensis ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = sa.otx_genes$orthogroup)+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))

sa_dotplot_otx_only.e <- DotPlot(sa.ephyra, features =sa.otx_only_genes$Sanderia_malayensis ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = sa.otx_only_genes$orthogroup)+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))


sa_dotplot_unch_homeo.e <-DotPlot(sa.ephyra, features =sa.unch_home_gene$Sanderia_malayensis ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = sa.unch_home_gene$orthogroup)+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))

sa_dotplot_otx_cluster.e
sa_dotplot_otx_only.e
sa_dotplot_unch_homeo.e

ggsave(
  filename = file.path(out_dir, "sa_dotplot_otx_cluster.png"),
  plot = sa_dotplot_otx_cluster.e,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "sa_dotplot_otx_only.png"),
  plot = sa_dotplot_otx_only.e,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "sa_dotplot_unch_homeo.png"),
  plot = sa_dotplot_unch_homeo.e,
  width = 10,
  height = 6,
  dpi = 300
)


```

```{r}
#CPATH
sa_dotplot_cpath.e <-DotPlot(sa.ephyra, features ="ENSXQXG00000001686" ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = gene_labels["ENSXQXG00000001686"])

sa_dotplot_cpath.e

ggsave(
  filename = file.path(out_dir, "sa_dotplot_cpath.png"),
  plot = sa_dotplot_cpath.e,
  width = 10,
  height = 6,
  dpi = 300
)

```


```{r}
#Hox1
sa_dotplot_hox1.e <-DotPlot(sa.ephyra, features ="ENSXQXG00000003349")+
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis()+scale_x_discrete(labels = gene_labels["ENSXQXG00000003349"])

sa_dotplot_hox1.e

ggsave(
  filename = file.path(out_dir, "sa_dotplot_hox1.png"),
  plot = sa_dotplot_hox1.e,
  width = 10,
  height = 6,
  dpi = 300
)


```



```{r}
#FoxB genes
sa.foxB_genes <- c('ENSXQXG00000008379','ENSXQXG00000008891','ENSXQXG00000011219')

sa_dotplot_foxb <- DotPlot(sa.ephyra, features =sa.foxB_genes ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = c('foxB1','foxB2','foxB3'))

ggsave(
  filename = file.path(out_dir, "sa_dotplot_foxb.png"),
  plot = sa_dotplot_foxb,
  width = 10,
  height = 6,
  dpi = 300
)


```

```{r}
#btbd2
sa_dotplot_btb <- DotPlot(sa.ephyra, features =c("ENSXQXG00000011094")) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis()+scale_x_discrete(labels = c('btbd2-like-11'))+
  theme(plot.title = element_text(hjust = 0.5))+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))

sa_dotplot_btb

ggsave(
  filename = file.path(out_dir, "sa_dotplot_btb.png"),
  plot = sa_dotplot_btb,
  width = 10,
  height = 6,
  dpi = 300
)

```

Showing that these candidates are not expressed in the polyp
```{r}
sa_dotplot_tf_candi_lifehistory <- DotPlot(data.sa, features = c("ENSXQXG00000001686", "ENSXQXG00000016532","ENSXQXG00000003349","ENSXQXG00000011094"), group.by = "orig.ident")+
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis()+scale_x_discrete(labels = gene_labels[c("ENSXQXG00000001686", "ENSXQXG00000016532","ENSXQXG00000003349","ENSXQXG00000011094")])+
     theme(axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ), axis.text.y =element_text(size=15, face="bold" ))
sa_dotplot_tf_candi_lifehistory

ggsave(
  filename = file.path(out_dir, "sa_dotplot_tf_candi_lifehistory.png"),
  plot = sa_dotplot_tf_candi_lifehistory,
  width = 8,
  height = 6,
  dpi = 300
)
```

#Biased approach

Plotting the biased approach transcription factors by importing the built orthology table
```{r}
#Importing the table based on the OMA table and OrthoSearch script 
bilat_tf_table <- read.table(file = "../../data/seqs/bilat_tfs/bilat_tf_orthology_table.tsv",header = T,fill = T)

#removing hlxb9 since it is not hlxB9 (not present in medusozoan)+all genes for which 
#no litterature could be found
filter_out <- c('hlxb9','HOX1','FEZF1','OTP','RAX','EMX')

bilat_tf_table <- bilat_tf_table[!(bilat_tf_table$gene %in% filter_out),] 
bilat_tf_sa <- bilat_tf_table$Sanderia_gene


# Build the named vector: names = original, values = display labels
bilat_tf_sa_label <- setNames(bilat_tf_table$gene, bilat_tf_table$Sanderia_gene)

bilat_tf_sa_plot <- DotPlot(
  sa.ephyra,
  features = bilat_tf_sa
) +
  RotatedAxis() +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  scale_x_discrete(labels = bilat_tf_sa_label)+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(angle = 90, vjust = 0.5, hjust = 1))
bilat_tf_sa_plot

ggsave(
  filename = file.path(out_dir, "sa_bilat_tfs_updated_dotplot.png"),
  plot = bilat_tf_sa_plot,
  width = 14,
  height = 9,
  dpi = 300
)

```




##Saving
Saving all the plots 

```{r}
out_dir <- '../../screenshots/R/Sa/final/'

ggsave(
  filename = file.path(out_dir, "sa_dimplot_coarse2.png"),
  plot = sa_dimplot_coarse,
  width = 14,
  height = 9,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "sa_dimplot_fine2.png"),
  plot = sa_dimplot_fine,
  width = 14,
  height = 9,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "sa_dimplot_p_v_e2.png"),
  plot = sa_dimplot_fine_p_v_e,
  width = 14,
  height = 9,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "sa_dotplot_STm_tf_markers_all.png"),
  plot = sa_dotplot_STm_tf_markers_all,
  width = 14,
  height = 9,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "sa_dotplot_STm_tf_markers_top10.png"),
  plot = sa_dotplot_STm_tf_markers_top10,
  width = 14,
  height = 9,
  dpi = 300
)

```

```{r}
write.table(sa.tf.muscleST_markers.e, file = "../../results/R/expressed_tfs_lists/sa_ST_expressed_tf.tsv", sep = "\t",  row.names = FALSE, quote = FALSE)
write.table(sa.tf.markers_fine.e, file = "../../results/R/expressed_tfs_lists/sa_tf_markers_fine_clusters.tsv", sep = "\t",  row.names = FALSE, quote = FALSE)
write.table(sa.all.markers_fine.e, file = "../../results/R/expressed_tfs_lists/sa_all_markers_fine_clusters.tsv", sep = "\t",  row.names = FALSE, quote = FALSE)

save(sa.all.markers_coarse.e, sa.all.markers_fine.e, sa.tf.markers_coarse.e, sa.tf.markers_fine.e, file ="../../data/scyphozoa/sa_markers.RData")

```

