---
title: "Aurelia single-cell analysis"
author: "Matéo Peyroche"
format: html
---


#DATA ANALYSIS OF THE AURELIA DATASET

##Importing libraries, dataset and relevant objects

Libraries
```{r}
#| echo: false
set.seed(213)
library(Seurat,quietly=T)
library(SeuratObject,quietly=T)
library(RColorBrewer,quietly=T)
library(Matrix,quietly=T)
library(ggplot2,quietly=T)
library(Polychrome,quietly=T)
library(dplyr,quietly=T)
library(presto,quietly=T)
library(rtracklayer,quietly=T)
library(stringr,quietly=T)
library(tidyverse,quietly=T)
library(colorspace,quietly=T)
library(patchwork,quietly=T)
library(scCustomize,quietly=T)


```

Loading the dataset and the gene annotations
```{r}
load (file= "../../data/Ac_manuscript_revision_ACOE/Ac.Alldata.Robj")
load (file= "../../data/scyphozoa_AGC/ACOE.genes.RData")

out_dir <- '../../screenshots/R/Ac/final/' #output directory for figures

data.ac <- Ac.Alldata
ac.ephyra <- subset(data.ac, subset = orig.ident %in% c("ephyra.early","ephyra.late"))

```

```{r}

Idents(data.ac) <- data.ac$IDs
Idents(ac.ephyra) <- ac.ephyra$IDs


#match the idents with the colors of the palette
coarse_cluster_levels <- levels(Idents(data.ac))


coarse_pal <- createPalette(
  seedcolors = c("#EE7733", "#009988", "#EE3377"),
  N = length(coarse_cluster_levels),
  range = c(20,90)
)

names(coarse_pal) <- coarse_cluster_levels
coarse_pal <- sortByHue(coarse_pal)


ac_dimplot_eph_coarse <- DimPlot(
  ac.ephyra,
  reduction = "umap.h",
  cols = coarse_pal,
  pt.size = 0.5
)


Idents(data.ac) <- data.ac$ID.separate
Idents(ac.ephyra) <- ac.ephyra$ID.separate

fine_cluster_levels <- levels(Idents(ac.ephyra))
fine_pal <- character(length(fine_cluster_levels))
names(fine_pal) <- fine_cluster_levels

tab <- table(fine = ac.ephyra@meta.data$ID.separate, coarse = ac.ephyra@meta.data$IDs)

# for each fine cluster (row), pick the coarse cluster (column) with the most cells =
#pair fine clusters to the coarse cluster they are derived from
fine_to_coarse <- apply(tab, 1, function(x) names(which.max(x)))

#Derive fine cluster color from the coarse one

for (coarse_id in names(coarse_pal)) {
  children <- names(fine_to_coarse)[fine_to_coarse == coarse_id]
  n_children <- length(children)
  base_col <- coarse_pal[coarse_id]

  if (n_children == 1) {
    # no subclustering -> keep the coarse color exactly as-is
    fine_pal[children] <- base_col
  } else {
    # spread shades from darker to lighter around the base color
    shades <- sapply(seq(-0.4, 0.4, length.out = n_children), function(amt) {
      if (amt < 0) darken(base_col, abs(amt)) else lighten(base_col, amt)
    })
    fine_pal[children] <- shades
  }
}

ac_dimplot_fine <-DimPlot(
  ac.ephyra,
  reduction = "umap.h",
  cols = fine_pal,
  pt.size = 0.5
)
ac_dimplot_eph_coarse
ac_dimplot_fine

ggsave(
  filename = file.path(out_dir, "ac_dimplot_eph_coarse.png"),
  plot = ac_dimplot_eph_coarse,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "ac_dimplot_eph_fine.png"),
  plot = ac_dimplot_fine,
  width = 12,
  height = 6,
  dpi = 300
)
```

Plotting of Aurelia striated muscle specific genes
```{r}
#Plotting stmyhc1 and 4 and the "calm-like-37" 
ac_dotplot_stm_struct_genes.e <- DotPlot(ac.ephyra, features = c('stmyhc1','stmyhc4','calm-like-37')) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis()

ac_featplot_stm_stmyhc.e <- FeaturePlot(ac.ephyra, features = c('stmyhc1','stmyhc4','calm-like-37'),reduction = "umap.h")&
  scale_color_gradientn(colors = c("lightgrey", "#FDBB84", "#B30000"))

ac_dotplot_stm_struct_genes.e
ac_featplot_stm_stmyhc.e

ggsave(
  filename = file.path(out_dir, "ac_dotplot_stm_struct_genes.png"),
  plot = ac_dotplot_stm_struct_genes.e,
  width = 12,
  height = 14,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "ac_featplot_stm_stmyhc.png"),
  plot = ac_featplot_stm_stmyhc.e,
  width = 14,
  height = 9,
  dpi = 300
)

```

#Unbiased approach

##Ephyra Differentially expressed genes (DEG)

As we are interested in the striated muscles, we will work on the ephyra.
Getting DEGs for both clustering and for the TF genes subset 
```{r}
Idents(ac.ephyra) <- ac.ephyra$IDs

ac.all.markers_coarse.e <- FindAllMarkers(
  ac.ephyra,
  assay = "RNA",
  logfc.threshold = 0.25,
  return.thresh   = 0.001,
  min.pct         = 0.2,
  only.pos        = TRUE
)

ac.tf.markers_coarse.e <- FindAllMarkers(
  ac.ephyra,
  features= intersect(AcTF_list,rownames(ac.ephyra)),
  assay = "RNA",
  logfc.threshold = 0.25,
  return.thresh   = 0.001,
  min.pct         = 0.06,
  only.pos        = TRUE
)

Idents(ac.ephyra) <- ac.ephyra$ID.separate

ac.all.markers_fine.e <- FindAllMarkers(
  ac.ephyra,
  assay = "RNA",
  logfc.threshold = 0.25,
  return.thresh   = 0.001,
  min.pct         = 0.2,
  only.pos        = TRUE
)

ac.tf.markers_fine.e <- FindAllMarkers(
  ac.ephyra,
  features= intersect(AcTF_list,rownames(ac.ephyra)),
  assay = "RNA",
  logfc.threshold = 0.25,
  return.thresh   = 0.001,
  min.pct         = 0.06,
  only.pos        = TRUE
)

```

Extracting the ST-muscle markers, only for the coarse clustering to get the list for all the striated muscles. 
```{r}
ac.muscleST_markers.e <- ac.all.markers_coarse.e %>%
  filter(cluster == "muscle.st") %>%
  arrange(p_val_adj,desc(avg_log2FC))

ac.tf.muscleST_markers.e <- ac.tf.markers_coarse.e %>%
  filter(cluster == "muscle.st") %>%
  arrange(p_val_adj,desc(avg_log2FC))
```


Plotting the differentially expressed transcription factors.
```{r}

Idents(ac.ephyra) <- ac.ephyra$IDs
ac_dotplot_STm_tf_markers_all_coarse <- DotPlot(ac.ephyra, features = ac.tf.muscleST_markers.e$gene) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis()+labs(title = "Differentially expressed transcription factors of the striated muscle cluster",size = 20, face = "bold")

print(ac_dotplot_STm_tf_markers_all_coarse)

ac_dotplot_STm_tf_markers_top10_coarse<- DotPlot(ac.ephyra, features = ac.tf.muscleST_markers.e[c(1:7,9:11),]$gene) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis()+labs(title = "Top differentially expressed transcription factors of the striated muscle cluster",size = 20, face = "bold")+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))

print(ac_dotplot_STm_tf_markers_top10_coarse)

Idents(ac.ephyra) <- ac.ephyra$ID.separate
ac_dotplot_STm_tf_markers_all_fine <- DotPlot(ac.ephyra, features = ac.tf.muscleST_markers.e$gene) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis()+labs(title = "Differentially expressed transcription factors of the striated muscle cluster",size = 20, face = "bold")

print(ac_dotplot_STm_tf_markers_all_fine)

ac_dotplot_STm_tf_markers_top10_fine<- DotPlot(ac.ephyra, features = ac.tf.muscleST_markers.e[c(1:7,9:11),]$gene) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis()+labs(title = "Top differentially expressed transcription factors of the striated muscle cluster",size = 20, face = "bold")+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))

print(ac_dotplot_STm_tf_markers_top10_fine)

ggsave(
  filename = file.path(out_dir, "ac_dotplot_STm_tf_markers_all_coarse.png"),
  plot = ac_dotplot_STm_tf_markers_all_coarse,
  width = 12,
  height = 9,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "ac_dotplot_STm_tf_markers_top10_coarse.png"),
  plot = ac_dotplot_STm_tf_markers_top10_coarse,
  width = 12,
  height = 9,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "ac_dotplot_STm_tf_markers_all_fine.png"),
  plot = ac_dotplot_STm_tf_markers_all_fine,
  width = 14,
  height = 14,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "ac_dotplot_STm_tf_markers_top10_fine.png"),
  plot = ac_dotplot_STm_tf_markers_top10_fine,
  width = 14,
  height = 14,
  dpi = 300
)

```

###Individual look at genes of interest
```{r}
#Otx genes 
#Importing the orthology table from the tree. It includes the uncharacterized homeobox that sits in the cluster
full_otx_ortho_table <- read_tsv("../../synteny/ortho_table_otx_synteny.tsv")
ac.otx_genes <- full_otx_ortho_table %>%
  select(orthogroup,Aurelia_coerulea)%>%
  filter (!is.na(Aurelia_coerulea))

ac.otx_genes <- ac.otx_genes %>%
  left_join(genes.ac %>% select(gene_short_name,cellranger), by = c('Aurelia_coerulea'='cellranger'))
ac.otx_genes$orthogroup[6] <- "unch_homeo"

ac.otx_only_genes <- ac.otx_genes[-6,]
ac.unch_home_gene <- ac.otx_genes[6,]



ac_dotplot_otx_cluster.e <- DotPlot(ac.ephyra, features =ac.otx_genes$gene_short_name ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = ac.otx_genes$orthogroup)+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))

ac_dotplot_otx_cluster.e.coarse <- DotPlot(ac.ephyra, features =ac.otx_genes$gene_short_name, group.by = "IDs" ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = ac.otx_genes$orthogroup)+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))


ac_dotplot_otx_only.e <- DotPlot(ac.ephyra, features =ac.otx_only_genes$gene_short_name ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = ac.otx_only_genes$orthogroup)+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))


ac_dotplot_unch_homeo.e <-DotPlot(ac.ephyra, features =ac.unch_home_gene$gene_short_name ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = ac.unch_home_gene$orthogroup)+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))

ac_dotplot_otx_cluster.e
ac_dotplot_otx_cluster.e.coarse
ac_dotplot_otx_only.e
ac_dotplot_unch_homeo.e

ggsave(
  filename = file.path(out_dir, "ac_dotplot_otx_cluster.png"),
  plot = ac_dotplot_otx_cluster.e,
  width = 12,
  height = 14,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "ac_dotplot_otx_cluster_coarse.png"),
  plot = ac_dotplot_otx_cluster.e.coarse,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "ac_dotplot_otx_only.png"),
  plot = ac_dotplot_otx_only.e,
  width = 12,
  height = 14,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "ac_dotplot_unch_homeo.png"),
  plot = ac_dotplot_unch_homeo.e,
  width = 12,
  height = 14,
  dpi = 300
)

```

```{r}
#CPATH
ac_dotplot_cpath.e <-DotPlot(ac.ephyra, features ="cPath" ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis()

ac_dotplot_cpath.e

ggsave(
  filename = file.path(out_dir, "ac_dotplot_cpath.png"),
  plot = ac_dotplot_cpath.e,
  width = 12,
  height = 14,
  dpi = 300
)

```

```{r}
#Hox1
ac_dotplot_hox1.e <-DotPlot(ac.ephyra, features ="hoxa1-like")+
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)

ac_dotplot_hox1.e

ggsave(
  filename = file.path(out_dir, "ac_dotplot_hox1.png"),
  plot = ac_dotplot_hox1.e,
  width = 12,
  height = 14,
  dpi = 300
)

```


```{r}
#FoxB genes
ac.foxB_genes <- c('foxb1-like-1','foxa1.1')

ac_dotplot_foxb <- DotPlot(ac.ephyra, features =ac.foxB_genes ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = c('foxB2','foxB3'))

ac_dotplot_foxb

ggsave(
  filename = file.path(out_dir, "ac_dotplot_foxb.png"),
  plot = ac_dotplot_foxb,
  width = 12,
  height = 14,
  dpi = 300
)

#Showing expression of foxB1 in a previous transcriptome-mapped object
load (file= '../../data/scyphozoa/aurelia.medusa.Robj')
Idents(aurelia) <- aurelia$IDs
ac_dotplot_foxb1 <- DotPlot(aurelia, features = "AaFoxB1-like" ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = c('foxB1'))+theme(axis.title = element_blank())
ac_dotplot_foxb1

ggsave(
  filename = file.path(out_dir, "ac_dotplot_foxb1.png"),
  plot = ac_dotplot_foxb1,
  width = 10,
  height = 6,
  dpi = 300
)

```

```{r}
#btbd2
ac_dotplot_btb <- DotPlot(ac.ephyra, features =c("btbd2-like-11")) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis()

ac_dotplot_btb

ggsave(
  filename = file.path(out_dir, "ac_dotplot_btb.png"),
  plot = ac_dotplot_btb,
  width = 12,
  height = 14,
  dpi = 300
)

```

Showing that these candidates are not expressed in the polyp
```{r}
ac_dotplot_tf_candi_lifehistory <- DotPlot(data.ac, features = c("cPath", "otx1a","hoxa1-like","btbd2-like-11"), group.by = "lifehistory")+
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis()+
     theme(axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ), axis.text.y =element_text(size=15, face="bold" ))
ac_dotplot_tf_candi_lifehistory

ggsave(
  filename = file.path(out_dir, "ac_dotplot_tf_candi_lifehistory.png"),
  plot = ac_dotplot_tf_candi_lifehistory,
  width = 8,
  height = 8,
  dpi = 300
)


```



##Coexpression of transcription factors and strucutural genes. 

```{r}
expr_data <- FetchData(ac.ephyra, vars = c("stmyhc5", "otx1a"))
#FIltering for cells with expression of the two genes to remove noise
cells_keep <- rownames(expr_data)[expr_data$stmyhc5 > 0 & expr_data$otx1a > 0.5]
data_filtered <- subset(ac.ephyra, cells = cells_keep)
ac_featscatter_otxE_stmyhc5 <- FeatureScatter(data_filtered, feature1 = "stmyhc5", feature2 = "otx1a", group.by = "IDs")
ac_featscatter_otxE_stmyhc5
ggsave(
  filename = file.path(out_dir, "ac_featscatter_otxE_stmyhc5.png"),
  plot = ac_featscatter_otxE_stmyhc5,
  width = 12,
  height = 8,
  dpi = 300
)

```



#Biased approach

Plotting the biased approach transcription factors by importing the built orthology table
```{r}
#Importing the table based on the OMA table and OrthoSearch script 
bilat_tf_table <- read.table(file = "../../data/seqs/bilat_tfs/bilat_tf_orthology_table.tsv",header = T,fill = T)

#removing hlxb9 since it is not hlxB9 (not present in medusozoan)+all genes for which 
#no litterature could be found
filter_out <- c('hlxb9','HOX1','FEZF1','OTP','RAX','EMX')

bilat_tf_table <- bilat_tf_table[!(bilat_tf_table$gene %in% filter_out),] 
bilat_tf_table <- bilat_tf_table %>%
  left_join(genes.ac %>% select(gene_short_name,cellranger), by = c('Aurelia_gene'='cellranger'))

bilat_tf_ac <- bilat_tf_table$gene_short_name


# Build the named vector: names = original, values = display labels
bilat_tf_ac_label <- setNames(bilat_tf_table$gene, bilat_tf_table$gene_short_name)

bilat_tf_ac_plot <- DotPlot(
  ac.ephyra,
  features = bilat_tf_ac
) +
  RotatedAxis() +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  scale_x_discrete(labels = bilat_tf_ac_label)+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(angle = 90, vjust = 0.5, hjust = 1))

bilat_tf_ac_plot

ggsave(
  filename = file.path(out_dir, "ac_bilat_tfs_updated_dotplot.png"),
  plot = bilat_tf_ac_plot,
  width = 16,
  height = 14,
  dpi = 300
)

```

##Saving
```{r}
write.table(ac.tf.muscleST_markers.e, file = "../../results/R/expressed_tfs_lists/ac_ST_expressed_tf.tsv", sep = "\t",  row.names = FALSE, quote = FALSE)
write.table(ac.tf.markers_fine.e, file = "../../results/R/expressed_tfs_lists/ac_tf_markers_fine_clusters.tsv", sep = "\t",  row.names = FALSE, quote = FALSE)
write.table(ac.all.markers_fine.e, file = "../../results/R/expressed_tfs_lists/ac_all_markers_fine_clusters.tsv", sep = "\t",  row.names = FALSE, quote = FALSE)

save(ac.all.markers_coarse.e, ac.all.markers_fine.e, ac.tf.markers_coarse.e, ac.tf.markers_fine.e, file ="../../data/scyphozoa/ac_markers.RData")

```


