---
title: "Clytia single-cell analysis"
author: "Matéo Peyroche"
format: html
---

#DATA ANALYSIS OF THE CLYTIA MEDUSA CAZET DATASET

##Importing libraries, dataset and relevant objects

Libraries
```{r}
#| echo: false
set.seed(213)
library(Seurat,quietly=T)
library(SeuratObject,quietly=T)
library(RColorBrewer,quietly=T)
library(Matrix,quietly=T)
library(ggplot2,quietly=T)
library(Polychrome,quietly=T)
library(dplyr,quietly=T)
library(presto,quietly=T)
library(rtracklayer,quietly=T)
library(stringr,quietly=T)
library(tidyverse,quietly=T)
library(colorspace,quietly=T)
library(patchwork,quietly=T)
library(scCustomize,quietly=T)
```

Loading the dataset and the gene annotations
```{r}
load (file= "../../data/scyphozoa_AGC/clytia.cazet.RData")
load (file= "../../data/scyphozoa_AGC/cazet.genes.annot.RData")
annots.ch <- read_tsv(file = "../../clytia/clytia_annotations/c_hem_nr_func_annotations.tsv")

#Getting the annotations from the latest gene set, smaller than the one used in the Chari et al. paper
#This is only to get names, we import the TF list for this set and not the newest one
genes.Ch <- genes.Ch[,-2]
annots.ch$gene <- gsub ('_','-', annots.ch$gene)
genes.Ch <- genes.Ch %>%
  left_join(annots.ch %>% select(gene, name),by=(c('gene'='gene')))

#Getting the gene column in the table and reformatting it to feat the gene names in the dataset
Ch_TF_table <- read.csv(file = '../../data/seqs/TF_lists/ClyHe_TF_List.csv')
Ch_TF_list <- Ch_TF_table [,c(1,2)]

Ch_TF_list$Gene_name <- gsub("^Chem[-_]", "", Ch_TF_list$Gene_name)
Ch_TF_list$Gene_name <- gsub("_", "-", Ch_TF_list$Gene_name)
Ch_TF_list <- unique(Ch_TF_list$Gene_name)

genes.Ch$gene <- sub("\\.mrna1$", "", genes.Ch$gene)
genes.Ch$gene <- sub("\\.mrna2$", "", genes.Ch$gene)

ch.medusa <- data1
out_dir <- '../../screenshots/R/Ch/final/' 

```

#Visualization of the dataset

```{r}
Idents(ch.medusa) <- ch.medusa$annosSub
cluster_levels <- levels(Idents(ch.medusa))
pal <- createPalette(
  seedcolors = c("#EE7733", "#009988", "#EE3377"),
  N = length(cluster_levels),
  range = c(20,90)
)
names(pal) <- cluster_levels

ch_dimplot <- DimPlot(
  ch.medusa,
  reduction = "umap",
  cols = pal,
  pt.size = 0.5
)

ch_dimplot

ch_highlight_plot_ST <- Cluster_Highlight_Plot(
  seurat_object = ch.medusa,
  reduction = "umap",
  cluster_name = c("Striated Muscle of Subumbrella", "Striated Muscle of Velum"),
  highlight_color = c("#E69F00","#0072B2"),
  background_color = "lightgray"
)

ch_highlight_plot_ST

ggsave(
  filename = file.path(out_dir, "ch_dimplot.png"),
  plot = ch_dimplot,
  width = 14,
  height = 9,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "cch_highlight_plot_ST.png"),
  plot = ch_highlight_plot_ST,
  width = 14,
  height = 9,
  dpi = 300
)

```


Creating gene labels for dotplots by importing names from the annotation file
```{r}
gene_labels <- genes.Ch %>%
  group_by(gene) %>%
  summarise(label = dplyr::first(name), .groups = "drop") %>%
  mutate(label = ifelse(nchar(label) > 20, paste0(substr(label, 1, 20), "..."), label)) %>%
  tibble::deframe()
```


Plotting of orthologs of Aurelia striated muscle specific genes

```{r}
oma.ChAc <- read_tsv(file = "../../data/OMA/ch-kostya_repaired.tsv")
load (file= "../../data/scyphozoa_AGC/ACOE.genes.RData")

ch_dotplot_stm_struct_genes <- DotPlot(ch.medusa, features = c("XLOC-029205","XLOC-038183","XLOC-041322")) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = c("st-myhc1.1","st-myhc1.2","st-myhc5"))

ch_featplot_stm_stmyhc <- FeaturePlot(ch.medusa, features = c("XLOC-029205","XLOC-038183","XLOC-041322"),reduction = "umap")&
  scale_color_gradientn(colors = c("lightgrey", "#FDBB84", "#B30000"))

ch_dotplot_stm_struct_genes
ch_featplot_stm_stmyhc

ggsave(
  filename = file.path(out_dir, "ch_dotplot_stm_struct_genes.png"),
  plot = ch_dotplot_stm_struct_genes,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "ch_featplot_stm_stmyhc.png"),
  plot = ch_featplot_stm_stmyhc,
  width = 14,
  height = 9,
  dpi = 300
)
```


#Unbiased approach

##Clytia medusa differentially expressed genes (DEG)

Getting the markers for both the whole set and the TFS

```{r}
#We need to resolve dodgy Trinity names
# Check current problematic names first
head(grep("\\.mrna", rownames(ch.medusa), value = TRUE))

new_names <- sub("\\.mrna\\d+$", "", rownames(ch.medusa))

# check for uniqueness before renaming
sum(duplicated(new_names))

#Now changing the rownames
counts_matrix <- GetAssayData(ch.medusa, assay = "RNA", layer = "counts")
rownames(counts_matrix) <- sub("\\.mrna\\d+$", "", rownames(counts_matrix))

data_matrix <- GetAssayData(ch.medusa, assay = "RNA", layer = "data")
rownames(data_matrix) <- sub("\\.mrna\\d+$", "", rownames(data_matrix))

# Rebuild the assay (Seurat v5 approach)
ch.medusa[["RNA"]] <- CreateAssay5Object(counts = counts_matrix, data = data_matrix)
ch.medusa <- JoinLayers(ch.medusa)

ch.all.markers <- FindAllMarkers(
  ch.medusa,
  assay = "RNA",
  logfc.threshold = 0.25,
  return.thresh   = 0.001,
  min.pct         = 0.2,
  only.pos        = TRUE
)

ch.tf.markers <- FindAllMarkers(
  ch.medusa,
  features= intersect(Ch_TF_list,rownames(ch.medusa)),
  assay = "RNA",
  logfc.threshold = 0.25,
  return.thresh   = 0.001,
  min.pct         = 0.06,
  only.pos        = TRUE
)

```
##Striated muscle clusters
Extracting the gene markers for the striated muscle
```{r}
ch.umbrellaST_tf_markers <- ch.tf.markers %>%
  filter(cluster == "Striated Muscle of Subumbrella") %>%
  arrange(p_val_adj,desc(avg_log2FC))

ch.velumST_tf_markers <- ch.tf.markers %>%
  filter(cluster == "Striated Muscle of Velum") %>%
  arrange(p_val_adj,desc(avg_log2FC))

ch.radialSM_tf_markers <- ch.tf.markers %>%
  filter(cluster == "Radial Smooth Muscles") %>%
  arrange(p_val_adj,desc(avg_log2FC))

```


```{r}
ch_dotplot_subumbrellaST_tfs <- DotPlot(ch.medusa, features = ch.umbrellaST_tf_markers$gene) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = gene_labels[ch.umbrellaST_tf_markers$gene])+labs(title = "Differentially expressed transcription factors of the striated subumbrella muscles",size = 20, face = "bold")+
  theme(plot.title = element_text(hjust = 0.5))+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))

ch_dotplot_subumbrellaST_tfs

ch_dotplot_subumbrellaST_tfs_top4 <- DotPlot(ch.medusa, features = ch.umbrellaST_tf_markers$gene[1:4]) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = gene_labels[ch.umbrellaST_tf_markers$gene[1:4]])+labs(title = "Differentially expressed transcription factors of the striated subumbrella muscles",size = 20, face = "bold")+
  theme(plot.title = element_text(hjust = 0.5))+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))

ch_dotplot_subumbrellaST_tfs_top4

ch_dotplot_velumST_tfs <- DotPlot(ch.medusa, features = ch.velumST_tf_markers$gene) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = gene_labels[ch.velumST_tf_markers$gene])+labs(title = "Differentially expressed transcription factors of the striated velum muscles",size = 20, face = "bold")+
  theme(plot.title = element_text(hjust = 0.5))+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))

ch_dotplot_velumST_tfs

ch_dotplot_radialSM_tfs <- DotPlot(ch.medusa, features = ch.radialSM_tf_markers$gene) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = gene_labels[ch.radialSM_tf_markers$gene])+labs(title = "Differentially expressed transcription factors of the radial smooth muscles",size = 20, face = "bold")+
  theme(plot.title = element_text(hjust = 0.5))+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))

ch_dotplot_radialSM_tfs


ggsave(
  filename = file.path(out_dir, "ch_dotplot_subumbrellaST_tfs.png"),
  plot = ch_dotplot_subumbrellaST_tfs,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "ch_dotplot_subumbrellaST_tfs_top4.png"),
  plot = ch_dotplot_subumbrellaST_tfs_top4,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "ch_dotplot_velumST_tfs.png"),
  plot = ch_dotplot_velumST_tfs,
  width = 10,
  height = 6,
  dpi = 300
)

ggsave(
  filename = file.path(out_dir, "ch_dotplot_radialSM_tfs.png"),
  plot = ch_dotplot_radialSM_tfs,
  width = 10,
  height = 6,
  dpi = 300
)


```


###Individual look at genes of interest

```{r}
#otx genes
ch.otx_genes <- c('XLOC-039817','XLOC-000267','XLOC-040405')

ch_dotplot_otx <- DotPlot(ch.medusa, features =ch.otx_genes ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = c('otxA','otxC','otxE'))+labs(title = "Clytia otx genes expression pattern",size = 20, face = "bold")+
  theme(plot.title = element_text(hjust = 0.5))+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(size=15, face="bold" ))

ch_dotplot_otx

ggsave(
  filename = file.path(out_dir, "ch_dotplot_otx.png"),
  plot = ch_dotplot_otx,
  width = 10,
  height = 6,
  dpi = 300
)

```


```{r}
#cPath
ch_dotplot_cpath <- DotPlot(ch.medusa, features ='XLOC-001472' ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = c('cPath'))+labs(title = "Clytia cPath expression pattern",size = 20, face = "bold")+
  theme(plot.title = element_text(hjust = 0.5))

ch_dotplot_cpath

ggsave(
  filename = file.path(out_dir, "ch_dotplot_cpath.png"),
  plot = ch_dotplot_cpath,
  width = 10,
  height = 6,
  dpi = 300
)
```

```{r}
#Hox1
ch_dotplot_hox1 <- DotPlot(ch.medusa, features ='XLOC-038841' ) +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  RotatedAxis() +
 scale_x_discrete(labels = c('hox1'))+labs(title = "Clytia hox1 expression pattern",size = 20, face = "bold")+
  theme(plot.title = element_text(hjust = 0.5))

ch_dotplot_hox1

ggsave(
  filename = file.path(out_dir, "ch_dotplot_hox1.png"),
  plot = ch_dotplot_hox1,
  width = 10,
  height = 6,
  dpi = 300
)
```

#Biased approach

Plotting the biased approach transcription factors by importing the built orthology table
```{r}
#Importing the table based on the OMA table and OrthoSearch script 
bilat_tf_table <- read.table(file = "../../data/seqs/bilat_tfs/bilat_tf_orthology_table.tsv",header = T,fill = T)

#removing hlxb9 since it is not hlxB9 (not present in medusozoan)+all genes for which 
#no litterature could be found
filter_out <- c('hlxb9','HOX1','FEZF1','OTP','RAX','EMX')

bilat_tf_table <- bilat_tf_table[!(bilat_tf_table$gene %in% filter_out),] 
bilat_tf_table$Clytia_gene_2021 <- gsub('_','-',bilat_tf_table$Clytia_gene_2021)
bilat_tf_ch <- bilat_tf_table$Clytia_gene_2021

# Build the named vector: names = original, values = display labels
bilat_tf_ch_label <- setNames(bilat_tf_table$gene, bilat_tf_table$Clytia_gene_2021)

bilat_tf_ch_plot <- DotPlot(
  ch.medusa,
  features = bilat_tf_ch
) +
  RotatedAxis() +
  scale_color_gradient2(low = "#2166AC", mid = "white", high = "#B2182B", midpoint = 0)+
  scale_x_discrete(labels = bilat_tf_ch_label)+
     theme(axis.line = element_blank(),axis.ticks = element_blank(),axis.title = element_blank(),axis.text.x =element_text(angle = 90, vjust = 0.5, hjust = 1))
bilat_tf_ch_plot

ggsave(
  filename = file.path(out_dir, "bilat_tf_ch_plot.png"),
  plot = bilat_tf_ch_plot,
  width = 14,
  height = 9,
  dpi = 300
)

```
